master thesis

### 2018-03-26 Paper - Latex/Paper - Skizze - v4.tex
Paper - restructure

% Parasitism in Eukaryota - Reconstruction of ancestral and unavailable extant states

% Lydia Buntrock^1; Jorrit Poelen^?; Berhard Renard^2; Emanuel Heitlinger^1,3;

% 1 - Institute for Biology, Molecular Parasitology, Humboldt University, Berlin, Germany
% 2 - Bioinformatics Unit (MF1), Department for Methods Development and Research Infrastructure, Robert Koch Institute, Berlin, Germany
% 3 - Research Group Ecology and Evolution of Molecular Parasite Host Interactions, Leibniz Institute for Zoo and Wildlife Research, Berlin, Germany

% ## Wenn du willst kannst du auch das IZW (3) an deinen Namen heften!
% ## Yorrit bitte das Abstract schicken und fragen ob er Co-Author sein will.

  Parasitism can be defined as an interaction between species in which one of the interaction 
    partners, the parasite, lives in or on the other, the host. The parasite draws food from its 
    host and harms it in the process. According to some estimates, above 50\% of all eukaryotes are 
    parasites. Nevertheless, it is difficult to obtain information whether a particular taxon is a 
    parasite computationally making it difficult to query large sets of taxa. \\

  Here we test in how far it is possible to use the open tree of life (OTL), a synthesis of 
    phylogenetic trees on a backbone taxonomy (resulting in unresolved nodes), to expand available 
    information via phylogenetic trait prediction. We use the Global Biotic Interactions (GloBI) 
    database to categorise 25,992 and 34,879 species as parasites and free-living, respectively and 
    predict states for over ~2.7 million (97.6\%) leaf nodes without state information. \\

  We estimate the accuracy of our maximum parsimony based predictions using cross-evaluation and 
    simulation at roughly 80\% overall, but strongly varying between clades. We describe this 
    variation across taxa as associated with available state and topology information. We compare 
    our results with several smaller scale studies, which used manual expert curation and conclude 
    that computationally inferred state changes largely agree in number and placement with those. In 
    clades in which available state information is biased (mostly towards parasites, e.g. in 
    Nematodes) phylogenetic prediction is bound to provide results contradicting conventional 
    wisdom. \\

  This represents, to our knowledge, the first comprehensive computational reconstruction of the 
    emergence of parasitism in eukaryotes. We argue that such an approach is necessary to allow 
    further incorporation of parasitism as an important trait in species interaction databases and 
    in individual studies on eukaryotes e.g. in the microbiome. \\

  ------------ \\

  This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
    Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  The challenge here is the size of the tree and the little information about it. \\
  Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
    tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
    \cite{Poelen2014} which we used, we could only collect 25,992 parasitic and 34,879 free-living 
    species. It follows that we have $\approx 2.4~\%$ state information. \\
  So far, especially small scale studies have been carried out or highly manual. In this scale, it 
    requires different data sources to be interconnected. \\
  We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
    using the R package \textit{Castor} \cite{Louca2017}. \\
  Nevertheless, the results are convincing and even though purely computational approach which did 
    not include human experts input, results coincide with prior knowledge. Also regarding the number 
    of events, our estimates coincide with previous results by human experts, e.g. the study by 
    Weinstein and Kuris \cite{Weinstein2016}. \\

  ------------ \\

  \anmerkungstext{Klassisch packt man keine Referenzen in Abstracts (bzw. wenn das meist als 
    Kurzreferenz also (Author et al., 2017). (Bernhard)} \\
  We have compared the results of some subtrees with known knowledge (Chordata, Nematoda, 
    Platyhelminthes and Apicomplexa) and, except for the Nematoda, the results looked very good. In 
    the case of Nematoda, the data situation is strongly shifted to the few parasites. \\
  We could partly compare our number of origins with the results of Sara B. Weinstein and Armand M. 
  Kuris from their article \textit{Independent origins of parasitism in Animalia}  
    and have come to a similar magnitude. They identified 223 parasitic origins in Metazoa and we 
    were able to estimate about 300 origins. \\

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %--------------------------------------------------------------------------------------------------- 
  \section{Motivation}
    The aim of this thesis is the application of maximum parsimony algorithms to non-binary 
      trees and very large data sets. In particular, the example find the 'Origins of Parasitism' 
      throughout the Eukaryota Tree of Life. \\
    For these large phylogenetic synthesis trees, however, ancestral state reconstruction has so far 
      only been done for Bacteria and Archaea for binary traits by Goberna and Verdú \cite{Goberna2015}.
      However, this differs from eukaryotes in the sense that complex traits such as parasitism depend
      on more than one gene. \\
    \todo{Warum ist das relevant? Was könnte man dann tun?}

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %--------------------------------------------------------------------------------------------------- 
  \section{Background}
    For about 50 years, people have been working on ancestral state reconstruction. The first paper 
      is by Camin and Sokal, who in 1965 were working on algorithms for discrete-state data 
      \cite{Camin1965}. Different methods have been developed and the question is which method is the 
      most suitable for the problem at hand: The ancestral state reconstruction for a huge 
      multifurcated tree with \colorbox{red}{binary/two discrete} states. \\
    Royer-Carenzi et al. distinguishes two major classes of ancestral state reconstruction methods: \\
    The first is maximum parsimony: explain the current state with the least number of state changes 
      between the child and its ancestor. \\
    The other class she presents describes modeling the character evolution as a stochastic process and 
      using the likelihoods to compute the possible ancestral character states. This is generally done 
      with a continuous time Markov model \cite{RoyerCarenzi2013}. \\

    Parsimony methods are simple methods in which you can not include information such as branch 
      length (development time of the species). This is also their main point of criticism compared to 
      the more complex likelihood based models. Since there are no branch lengths or other information 
      available for the present relationship tree, and we thus can not take advantage of the 
      other models, we decided in favor for parsimonious. \\

    We will test the existing maximum parsimony algorithms of Fitch \cite{Fitch1971} and Sankoff 
      \cite{Sankoff1975} for this task and estimate their predictive power. The present tree structure 
      of OTL is not binary. A tree is multifurcated if each node has multiple ($n >= 3$) children 
      \cite{Felsenstein2003}. \\
    Maximum parsimony methods are developed for phylogenies, which are usually depicted as binary trees.
      Parsimonious in phylogeny refers to favoring the tree that needs the least evolutionary change 
      to explain the observed data. In our case, it is about the change of states 'is free-living' or 
      'is parasitic'. \\
      % Parsimonious in this context means to minimize the number of transitions between species states 
      % (free-living or parasitic). 
    The Sankoff method is implemented by Louca et al. for the non-binary case and is available as an R 
      package called \textit{Castor} \cite{Louca2017}. In addition, we have implemented the Fitch 
      method and adapted it for multifurcated trees. \\% to make it usable and comparable for our current problem.
    This achieves that we can predict ancestral states and unknown states of living species for large 
      non-binary relatives trees. \\

@@

    To accomplish this task, a large phylogenetic tree and information about the current species states 
      is needed. \\
    The biggest 'phylogenetic tree' is a synthesis of phylogenetic trees filled with a taxonomic trees 
      given by Open Tree of Life (OTL) \cite{Hinchliff2015}. %This tree is not binary and therefore the 
      % developed algorithms are not directly applicable. Therefore, an algorithm must be found that can 
      % work with non-binary trees. \\
    For the information about the current states of the species we use the interaction database Global 
      Biotic Interactions (GloBI) \cite{Poelen2014}. The data in GloBI are stored as interactions e.g. 
      species A parasitize species B. From this we conclude that species A is parasitic and species B 
      free-living. \\

    At this point a few words to the term parasitic. There are different definitions. Since we use 
      GloBI to classify species, we use their definition of parasitism. Again, in GloBi, Ontobee 
      definitions are used \cite{Xiang2011}.
    The interaction \textit{has parasite} is defined as: "An interaction relationship between two 
      organisms living together in more or less intimate association in a relationship in which 
      association is disadvantageous or destructive to one of the organisms."\footnote{
        \hyperlink{http://www.ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO_0002445}
        {ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO\_000244}; Last checked: 22.03.2018.
      }. This definition includes: ecto- and endoparasites, parasitoids, kleptoparasites and pathogenes.

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %--------------------------------------------------------------------------------------------------- 
  \section{Structure of this work}
    The objectives of this work are the following points: (1) Find a suitable ancestral state 
      reconstruction method. (2) Perform reconstructing on the Eucaryotic synthesis tree of OTL. \\
    The goal of Point 1 is to evaluate the possible methods based on a simulation of our data 
      situation. \\
    The one by Louca et al. implemented Sankoff algorithm is the best in our comparisons. Therefore, 
      point 2 consists of reconstructing the ancestral states and predicting the unknown leaf states. 
      And then perform an evaluation of the results.

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Description of data sets}
    Two types of data are needed for an ancestral state reconstruction: a tree and information about 
      the states. \\
    For the tree Open Tree of Life (OTL) is used \cite{Hinchliff2015} and for the state information 
      the Global Biotic Interaction database (GloBI) is taken \cite{Poelen2014}. \\
    % These are presented below.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{OTL}
      For this project a large database for phylogenetic trees and also for a taxonomic tree is needed. 
        Since an ancestral state reconstruction algorithm is applied to the phylogenetic tree, and for 
        the assessment and other properties the taxonomy provides much more information. \\
      OTL gives a synthesis of phylogenetic trees (currently 819 trees) and a taxonomic tree\footnote{
          \hyperlink{https://tree.opentreeoflife.org/about/synthesis-release/v9.1}
          {https://tree.opentreeoflife.org/about/synthesis-release/v9.1}; Last checked: 22.03.2018.
        }. OTL also 
        includes the large phylogenetic database TreeBASE \cite{Hinchliff2015}. \\
      \todo{Das steht auf der Website nicht in dem Paper...} \\
      For phylogenetic data, there are at least five big data collections, namely:
      \begin{itemize}
        \item ITIS (Integrated Taxonomic Information System) \cite{ITIS}
        \item NCBI (National Center for Biotechnology Information) \cite{NCBI1988}
        \item WORMS (World Register of Marine Species) \cite{WoRMS2018}
        \item GBIF (Global Biodiversity Information Facility) \cite{GBIF}
        \item OTT (OpenTreeOfLife-Taxonomy) \cite{Hinchliff2015}
      \end{itemize}
      ITIS is only a small set of 100~\% confirmed and named species. GBIF is not composed with the help 
        of phylogeny, the same is valid for the NCBI taxonomy. The WORMS taxonomy is a way too small 
        dataset of mostly marine species. \\
      Here the taxonomy from OTL is used because it is including most of the known taxonomies and is 
        synthesised by preffering taxonomies that match with available phylogenetic data. Furthermore 
        the team from OTL preferre a maximum number of species \cite{Hinchliff2015}. This is resulting 
        in somekind of hybrid between taxonomy and phylogeny. \anmerkungstext{Wie genau ist das ein 
        Hybrid? Genauer beschreiben, was Du damit meinst... (Thilo)} \\

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Transition probabilities}
      As mentioned above, in an ancestral state reconstruction, transitions are minimized, with 
        transition probabilities playing a role. \\
      This subsection deals with these transition probabilities from free-living (hereinafter / as a 
        formula FL) to parasitic (hereinafter P) and vice versa: $\mathcal{P}(FL \rightarrow P)$, 
        $\mathcal{P}(P \rightarrow FL)$. \\
      Different parasite types have different transition probabilities. It is very difficult to make a 
        statement about these probabilities. It is generally assumed in this work that there are 40~\% 
        parasites and 60~\% free-livings which is based on the estimates by Windsor \cite{Windsor1998} 
        and $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$, as a result of the 
        diversity of parasites and the lack of general determinations for this. These parameters are 
        debated in subsection \ref{sec:discussion - transition probabilities} of the discussion. \\

      For the maximum parsimony analysis of the real data, all transition probabilities are equated.
        However, the used castor package \cite{Louca2017} offers the possibility to enter different 
        transition probabilities.
      \begin{wrapfigure}{r}{0.5\textwidth}
        \begin{center}
          \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.45\textwidth]{Figures/40-60.png}
        \end{center}
        \caption{Simulated distributions of the states: \\
          60~\% Free-living - 40~\% Parasites \\
          red: parasites, blue: free-living, \\
          the threshold is at 0.4}
        \label{fig:Beta distribution}
      \end{wrapfigure}

      In the simulation two beta distributions have been chosen and a threshold that indicates the change 
        between states. \\
      Different thresholds with different beta distributions are simulated, with different distributions 
        of parasites and free-livings:
        \begin{itemize}
          \item 50~\% P to 50~\% FL,
          \item 40~\% P to 60~\% FL,
          \item 30~\% P to 70~\% FL and 
          \item 20~\% P to 80~\% FL
        \end{itemize}
        (\todo{see results simulation, ref...}). Figure \ref{fig:Beta distribution} 
        shows one example of these. \\

@@

        Each internal node is formed from its child nodes, where at the beginning the only information
        lies. \\
      Second traverse from the root node to the leafs. Each internal node is formed from its father node 
        and its sibling node. \\
      Last traversion (direction does not matter): Build the final state for every node. It is formed 
        from the sets of previous traversals. \\
      (The original Fitch algorithm was designed to minimize transitions without predicting actual states 
        of internal nodes, so it was just the first traversal.) \\
      The extension to the non-binary case is quite obvious, but holds some opportunities. In this case, 
        more than two children may be present for the first traversal, but the incision or union may 
        also be formed over more than two sets. Also in the second traversing, there may be several 
        sibling nodes. However, there are several possibilities here that were all tested and compared 
        in the simulation. Some of these options are already available in the binary case:
      \begin{itemize}
        \item The father node has (except for the root node) two state sets, because he came through 
          the up-traversing previously. Are both sets used or only the first traversing?
        \item Since there are several siblings, do \colorbox{red}{you} first of all make the cut or union, 
          or directly in the whole with the father node?
      \end{itemize}
      The first point already has an effect on the binary case. Figure \ref{fig: binary Fitch} shows 
        both possibilities of the three sets. \\
      Cunningham uses only the first state set of the father node \cite{Cunningham1998}. \\
      From these two points four different versions of Fitch were formed:
      \begin{enumerate}
        \item Fitch 1: First state set of father node; intersection/union of siblings first.
        \item Fitch 2: First state set of father node; intersection/union of siblings together with father node.
        \item Fitch 3: Both state sets of father node; intersection/union of siblings first.
        \item Fitch 4: Both state sets of father node; intersection/union of siblings together with father node sets.
      \end{enumerate}
      \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{Figures/simulation_fitch_evaluation.png}
        \caption{Test of Fitch Versions.}
        \label{fig:Fitch versions}
      \end{figure}
      These four versions were tested in the simulation with 100 trees and 10000 leaf nodes and a
        distribution of 60~\% FL to 40~\% P. Figure \ref{fig:Fitch versions} shows this over all unknown 
        node percentage. \\
      % At 90~\% unknown nodes and 90~\% of multifurcation of the internal nodes, version 1 was 89.26~\%, 
      %   version 2 was 89.26~\%, version 3 was 89.35~\%, and version 4 was 89.31~\% correct. Therefore, 
      %   only version 3 was used for all further simulations.
      At 95~\% unknown nodes and 95~\% of multifurcation of the internal nodes, version 1 was 88.37~\%, 
        version 2 was 88.37~\%, version 3 was 88.4~\%, and version 4 was 88.39~\% correct. Therefore, 
        only version 3 was used for all further simulations.

@@

  We then evaluated the results in two ways. On the one hand, we have examined the results of some 
    subtrees for their credibility with the help of biological background knowledge (subsection 
    \ref{sec:results - biological view}). On the other hand, we performed a leave-100-out cross-validation
    (Subsection \ref{sec:results - cross-validation}) and predicted approximately 98.17~\% states 
    correct.

@@

      Next to the problem of the multifurcation of the tree is the less of interaction data we have for 
        the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 distinct free-living species and 47,332 
        distinct parasite species for the whole tree of life. From the Eukaryota we could determine 
        25,992 and 34,879 species as parasites and free-living. With 2,293,463 leaf nodes we still have 
        about 97.34~\% unknown leaf nodes. In the discussion, section 
        \ref{sec:discussion - data situation}, we will talk about ways to extend this data. \\

@@

      Most species of Platyhelminthes (flatworms) are parasites, although there are also free-living, 
        predatory feeding species. These are summarized in the Turbellaria, while the parasites are 
        divided into three other classes \cite{Ax1961}. This also corresponds to our observations. There 
        is one class (Rhabditophora) that contains all but one single exception of free-living species 
        of this phylum, which includes the Turbellaria. \\
      It should be noted, however, that this classification is outdated, as it has been proven that 
        the Turbellaria are not monophyletic. But we will not go into that here. \\
      For the Platyhelminthes we had more state information for the leaf nodes compared to the other 
        considered subtrees, see Table \ref{table:percentage loss information subtrees}. We start with 
        0.1~\% free-livings and predicted 0.77~\% as free-living species. \\

>>>>

  Parasitism can be defined as an interaction between species in which one of the interaction 
    partners, the parasite, lives in or on the other, the host. The parasite draws food from its 
    host and harms it in the process. According to some estimates, above 40\% of all eukaryotes are 
    parasites. Nevertheless, it is difficult to obtain information whether a particular taxon is a 
    parasite computationally making it difficult to query large sets of taxa.

  Here we test in how far it is possible to use the Open Tree of Life (OTL), a synthesis of 
    phylogenetic trees on a backbone taxonomy (resulting in unresolved nodes), to expand available 
    information via phylogenetic trait prediction. We use the Global Biotic Interactions (GloBI) 
    database to categorise 25,992 and 34,879 species as parasites and free-living, respectively and 
    predict states for over $\sim$2.3 million (97.34\%) leaf nodes without state information.

  We estimate the accuracy of our maximum parsimony based predictions using cross-validation and 
    simulation at roughly 60\% overall, but strongly varying between clades. The cross-validation 
    resulted in an accuracy of 98.17\% which is explained by the fact that the data are not 
    uniformly distributed. We describe this variation across taxa as associated with available state 
    and topology information. We compare our results with several smaller scale studies, which used 
    manual expert curation and conclude that computationally inferred state changes largely agree in 
    number and placement with those. In clades in which available state information is biased 
    (mostly towards parasites, e.g. in Nematodes) phylogenetic prediction is bound to provide 
    results contradicting conventional wisdom.

  This represents, to our knowledge, the first comprehensive computational reconstruction of the 
    emergence of parasitism in eukaryotes. We argue that such an approach is necessary to allow 
    further incorporation of parasitism as an important trait in species interaction databases and 
    in individual studies on eukaryotes e.g. in the microbiome.

  % This study focuses on the ancestral state reconstruction of parasitism in the tree of life of
  %   Eukaryota. We predict unknown states of species and estimate origins and losses of parasitism. \\
  % The challenge here is the size of the tree and the little information about it. \\
  % Such a large phylogenetic tree does not completely exist and therefore we work with a synthesis 
  %   tree of OTL \cite{Hinchliff2015} which is highly multifurcated. \\
  % For the 2,535,437 leaf nodes we could not gather much data. From the GloBI database 
  %   \cite{Poelen2014} which we used, we could only collect 25,992 parasitic and 34,879 free-living 
  %   species. It follows that we have $\approx 2.4\%$ state information. \\
  % So far, especially small scale studies have been carried out or highly manual. In this scale, it 
  %   requires different data sources to be interconnected. \\
  % We performed an analysis of existing algorithms and selected a Sankoff maximum parsimony algorithm 
  %   using the R package \textit{Castor} \cite{Louca2017}. \\
  % Nevertheless, the results are convincing and even though purely computational approach which did 
  %   not include human experts input, results coincide with prior knowledge. Also regarding the number 
  %   of events, our estimates coincide with previous results by human experts, e.g. the study by 
  %   Weinstein and Kuris \cite{Weinstein2016}. \\

@@

  For about 50 years, people have been working on ancestral state reconstruction. The first paper 
    is by Camin and Sokal, who in 1965 were working on algorithms for discrete-state data 
    \cite{Camin1965}. Different methods have been developed and the question is which method is the 
    most suitable for the problem at hand: The ancestral state reconstruction for a huge 
    multifurcated tree with \colorbox{red}{binary/two discrete} states. \\
  Royer-Carenzi et al. distinguishes two major classes of ancestral state reconstruction methods: \\
  The first is maximum parsimony: explain the current state with the least number of state changes 
    between the child and its ancestor. \\
  The other class she presents describes modeling the character evolution as a stochastic process and 
    using the likelihoods to compute the possible ancestral character states. This is generally done 
    with a continuous time Markov model \cite{RoyerCarenzi2013}.

  Parsimony methods are simple methods in which you can not include information such as branch 
    length (development time of the species). This is also their main point of criticism compared to 
    the more complex likelihood based models. Since there are no branch lengths or other information 
    available for the present relationship tree, and we thus can not take advantage of the 
    other models, we decided in favor for parsimonious.

  We will test the existing maximum parsimony algorithms of Fitch \cite{Fitch1971} and Sankoff 
    \cite{Sankoff1975} for this task and estimate their predictive power. The present tree structure 
    of OTL is not binary. A tree is multifurcated if each node has multiple ($n >= 3$) children 
    \cite{Felsenstein2003}. \\
  Maximum parsimony methods are developed for phylogenies, which are usually depicted as binary trees.
    Parsimonious in phylogeny refers to favoring the tree that needs the least evolutionary change 
    to explain the observed data. In our case, it is about the change of states 'is free-living' or 
    'is parasitic'. \\
    % Parsimonious in this context means to minimize the number of transitions between species states 
    % (free-living or parasitic). 
  The Sankoff method is implemented by Louca et al. for the non-binary case and is available as an R 
    package called \textit{Castor} \cite{Louca2017}. In addition, we have implemented the Fitch 
    method and adapted it for multifurcated trees. \\% to make it usable and comparable for our current problem.
  This achieves that we can predict ancestral states and unknown states of living species for large 
    non-binary relatives trees.

@@

  To accomplish this task, a large phylogenetic tree and information about the current species states 
    is needed. \\
  The biggest 'phylogenetic tree' is a synthesis of phylogenetic trees filled with a taxonomic trees 
    given by Open Tree of Life (OTL) \cite{Hinchliff2015}. %This tree is not binary and therefore the 
    % developed algorithms are not directly applicable. Therefore, an algorithm must be found that can 
    % work with non-binary trees. \\
  For the information about the current states of the species we use the interaction database Global 
    Biotic Interactions (GloBI) \cite{Poelen2014}. The data in GloBI are stored as interactions e.g. 
    species A parasitize species B. From this we conclude that species A is parasitic and species B 
    free-living.

  At this point a few words to the term parasitic. There are different definitions. Since we use 
    GloBI to classify species, we use their definition of parasitism. Again, in GloBi, Ontobee 
    definitions are used \cite{Xiang2011}.
  The interaction \textit{has parasite} is defined as: "An interaction relationship between two 
    organisms living together in more or less intimate association in a relationship in which 
    association is disadvantageous or destructive to one of the organisms."\footnote{
      \hyperlink{http://www.ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO_0002445}
      {ontobee.org/ontology/RO?iri=http://purl.obolibrary.org/obo/RO\_000244}; Last checked: 22.03.2018.
    }. This definition includes: ecto- and endoparasites, parasitoids, kleptoparasites and pathogenes. \\

  The objectives of this work are the following points: (1) Find a suitable ancestral state 
    reconstruction method. (2) Perform reconstructing on the Eucaryotic synthesis tree of OTL. \\
  The goal of Point 1 is to evaluate the possible methods based on a simulation of our data 
    situation. \\
  The one by Louca et al. implemented Sankoff algorithm is the best in our comparisons. Therefore, 
    point 2 consists of reconstructing the ancestral states and predicting the unknown leaf states. 
    And then perform an evaluation of the results.

%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------------------
%--------------------------------------------------------------------------------------------------- 
\chapter{Aims}
  The aim of this thesis is the application of maximum parsimony algorithms to non-binary 
    trees and very large data sets. In particular, the example find the 'Origins of Parasitism' 
    throughout the Eukaryota Tree of Life. \\
  For these large phylogenetic synthesis trees, however, ancestral state reconstruction has so far 
    only been done for Bacteria and Archaea for binary traits by Goberna and Verdú \cite{Goberna2015}.
    However, this differs from eukaryotes in the sense that complex traits such as parasitism depend
    on more than one gene. \\
  \todo{Warum ist das relevant? Was könnte man dann tun?}

@@

  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  %---------------------------------------------------------------------------------------------------
  \section{Description of data sets}
    Two types of data are needed for an ancestral state reconstruction: a tree and information about 
      the states. \\
    For the tree Open Tree of Life (OTL) is used \cite{Hinchliff2015} and for the state information 
      the Global Biotic Interactions database (GloBI) is taken \cite{Poelen2014}. \\
    % These are presented below.

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{OTL}
      For this project a large database for phylogenetic trees and also for a taxonomic tree is needed. 
        Since an ancestral state reconstruction algorithm is applied to the phylogenetic tree, and for 
        the assessment and other properties the taxonomy provides much more information. \\
      OTL gives a synthesis of phylogenetic trees (currently 819 trees) and a taxonomic tree\footnote{
          \hyperlink{https://tree.opentreeoflife.org/about/synthesis-release/v9.1}
          {https://tree.opentreeoflife.org/about/synthesis-release/v9.1}; Last checked: 22.03.2018.
        }. OTL also 
        includes the large phylogenetic database TreeBASE \cite{Hinchliff2015}. \\
      \todo{Das steht auf der Website nicht in dem Paper...} \\
      For phylogenetic data, there are at least five big data collections, namely:
      \begin{itemize}
        \item ITIS (Integrated Taxonomic Information System) \cite{ITIS}
        \item NCBI (National Center for Biotechnology Information) \cite{NCBI1988}
        \item WORMS (World Register of Marine Species) \cite{WoRMS2018}
        \item GBIF (Global Biodiversity Information Facility) \cite{GBIF}
        \item OTT (OpenTreeOfLife-Taxonomy) \cite{Hinchliff2015}
      \end{itemize}
      ITIS is only a small set of 100\% confirmed and named species. GBIF is not composed with the help 
        of phylogeny, the same is valid for the NCBI taxonomy. The WORMS taxonomy is a way too small 
        dataset of mostly marine species. \\
      Here the taxonomy from OTL is used because it is including most of the known taxonomies and is 
        synthesised by preffering taxonomies that match with available phylogenetic data. Furthermore 
        the team from OTL preferre a maximum number of species \cite{Hinchliff2015}. This is resulting 
        in somekind of hybrid between taxonomy and phylogeny. \anmerkungstext{Wie genau ist das ein 
        Hybrid? Genauer beschreiben, was Du damit meinst... (Thilo)} \\

@@

    %---------------------------------------------------------------------------------------------------
    %---------------------------------------------------------------------------------------------------
    \subsection{Transition probabilities}
      As mentioned above, in an ancestral state reconstruction, transitions are minimized, with 
        transition probabilities playing a role. \\
      This subsection deals with these transition probabilities from free-living (hereinafter / as a 
        formula FL) to parasitic (hereinafter P) and vice versa: $\mathcal{P}(FL \rightarrow P)$, 
        $\mathcal{P}(P \rightarrow FL)$. \\
      Different parasite types have different transition probabilities. It is very difficult to make a 
        statement about these probabilities. It is generally assumed in this work that there are 40\% 
        parasites and 60\% free-livings which is based on the estimates by Windsor \cite{Windsor1998} 
        and $\mathcal{P}(FL \rightarrow P) = \mathcal{P}(P \rightarrow FL)$, as a result of the 
        diversity of parasites and the lack of general determinations for this. These parameters are 
        debated in subsection \ref{sec:discussion - transition probabilities} of the discussion. \\

      For the maximum parsimony analysis of the real data, all transition probabilities are equated.
        However, the used castor package \cite{Louca2017} offers the possibility to enter different 
        transition probabilities.
      \begin{wrapfigure}{r}{0.5\textwidth}
        \begin{center}
          \includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=0.45\textwidth]{Figures/40-60.png}
        \end{center}
        \caption{Simulated distributions of the states: \\
          60\% Free-living - 40\% Parasites \\
          red: parasites, blue: free-living, \\
          the threshold is at 0.4}
        \label{fig:Beta distribution}
      \end{wrapfigure}

      In the simulation two beta distributions have been chosen and a threshold that indicates the change 
        between states. \\
      Different thresholds with different beta distributions are simulated, with different distributions 
        of parasites and free-livings:
        \begin{itemize}
          \item 50\% P to 50\% FL,
          \item 40\% P to 60\% FL,
          \item 30\% P to 70\% FL and 
          \item 20\% P to 80\% FL
        \end{itemize}
        (\todo{see results simulation, ref...}). Figure \ref{fig:Beta distribution} 
        shows one example of these. \\

@@

        Each internal node is formed from its child nodes, where at the beginning the only information
        lies. \\
      Second traverse from the root node to the leafs. Each internal node is formed from its father node 
        and its sibling node. \\
      Last traversion (direction does not matter): Build the final state for every node. It is formed 
        from the sets of previous traversals. \\
      (The original Fitch algorithm was designed to minimize transitions without predicting actual states 
        of internal nodes, so it was just the first traversal.) \\
      The extension to the non-binary case is quite obvious, but holds some opportunities. In this case, 
        more than two children may be present for the first traversal, but the incision or union may 
        also be formed over more than two sets. Also in the second traversing, there may be several 
        sibling nodes. However, there are several possibilities here that were all tested and compared 
        in the simulation. Some of these options are already available in the binary case:
      \begin{itemize}
        \item The father node has (except for the root node) two state sets, because he came through 
          the up-traversing previously. Are both sets used or only the first traversing?
        \item Since there are several siblings, do \colorbox{red}{you} first of all make the cut or union, 
          or directly in the whole with the father node?
      \end{itemize}
      The first point already has an effect on the binary case. Figure \ref{fig: binary Fitch} shows 
        both possibilities of the three sets. \\
      Cunningham uses only the first state set of the father node \cite{Cunningham1998}. \\
      From these two points four different versions of Fitch were formed:
      \begin{enumerate}
        \item Fitch 1: First state set of father node; intersection/union of siblings first.
        \item Fitch 2: First state set of father node; intersection/union of siblings together with father node.
        \item Fitch 3: Both state sets of father node; intersection/union of siblings first.
        \item Fitch 4: Both state sets of father node; intersection/union of siblings together with father node sets.
      \end{enumerate}
      \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{Figures/simulation_fitch_evaluation.png}
        \caption{Test of Fitch Versions.}
        \label{fig:Fitch versions}
      \end{figure}
      These four versions were tested in the simulation with 100 trees and 10000 leaf nodes and a
        distribution of 60\% FL to 40\% P. Figure \ref{fig:Fitch versions} shows this over all unknown 
        node percentage. \\
      % At 90\% unknown nodes and 90\% of multifurcation of the internal nodes, version 1 was 89.26\%, 
      %   version 2 was 89.26\%, version 3 was 89.35\%, and version 4 was 89.31\% correct. Therefore, 
      %   only version 3 was used for all further simulations.
      At 95\% unknown nodes and 95\% of multifurcation of the internal nodes, version 1 was 88.37\%, 
        version 2 was 88.37\%, version 3 was 88.4\%, and version 4 was 88.39\% correct. Therefore, 
        only version 3 was used for all further simulations.

@@

  We then evaluated the results in two ways. On the one hand, we have examined the results of some 
    subtrees for their credibility with the help of biological background knowledge (subsection 
    \ref{sec:results - biological view}). On the other hand, we performed a leave-100-out cross-validation
    (Subsection \ref{sec:results - cross-validation}) and predicted approximately 98.17\% states 
    correct.

@@

      Next to the problem of the multifurcation of the tree is the less of interaction data we have for 
        the species. For the ancestral state reconstruction, we need information about the states 
        (free-living or parasite) in the leaf nodes. \\
      The eukaryotic synthesis tree has 293,463 leaf nodes. The GloBI database has 5,346,414 interactions 
        (at 29.01.2018). Out of this data we got 51,337 distinct free-living species and 47,332 
        distinct parasite species for the whole tree of life. From the Eukaryota we could determine 
        25,992 and 34,879 species as parasites and free-living. With 2,293,463 leaf nodes we still have 
        about 97.34\% unknown leaf nodes. In the discussion, section 
        \ref{sec:discussion - data situation}, we will talk about ways to extend this data. \\

@@

      Most species of Platyhelminthes (flatworms) are parasites, although there are also free-living, 
        predatory feeding species. These are summarized in the Turbellaria, while the parasites are 
        divided into three other classes \cite{Ax1961}. This also corresponds to our observations. There 
        is one class (Rhabditophora) that contains all but one single exception of free-living species 
        of this phylum, which includes the Turbellaria. \\
      It should be noted, however, that this classification is outdated, as it has been proven that 
        the Turbellaria are not monophyletic. But we will not go into that here. \\
      For the Platyhelminthes we had more state information for the leaf nodes compared to the other 
        considered subtrees, see Table \ref{table:percentage loss information subtrees}. We start with 
        0.1\% free-livings and predicted 0.77\% as free-living species. \\

      With the Nematoda it looks more complicated. Large parts of Nematoda are free-living, but we 
        found only 5.32\% of them. Blaxter et al. estimates the order of 25,000 parasites in the 
        Nematoda \cite{Blaxter2015} and speaks of at least seven independently arosed parasitism 
        \cite{Blaxter1998}. In a recent article Blaxter identifies 18 origins \cite{Blaxter2015} in 
        Nematoda. \\
      The problem at this point, however, is: Hallan speaks of the fact that only 23,000 species 
        were described by the Nematoda but is it assumed that one million or more species are in this 
        phylum.\footnote{J. Hallan, unpublished; 
          \hyperlink{https://insects.tamu.edu/research/collection/hallan/}
          {https://insects.tamu.edu/research/collection/hallan/}
        }. \todo{Link ist nicht erreichbar!} The parasites have been much more studied and thus we start 
        with only 0.63\% (table \ref{table:phylum leaf nodes states}) free-living species. Against such 
        a shifted data situation, the algorithm is almost powerless to make correct predictions. And yet 
        the percentage has increased to 5.34\%. \\
